Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-08-18 3dpas_intro.tex
small edits to applications work from Shantenu

The term ``application'' is often overloaded in literature and the computational
science discourse.  Sometimes the term is a reference to a stand-alone,
independent executable (e.g., a Molecular Dynamics simulation kernel such as
AMBER or Gromacs). Sometimes an application refers to science problem to which
the executable is put to use (e.g., compute the free-energy of binding energy of
a drug candidate), but sometimes it is also used to reference the end-to-end
workflow (i.e., a multi-stage computational process requiring a sequence of
possibly distinct simulations and executables).  To add chaos to the confusion,
applications are sometimes also as a proxy for projects (e.g., the ATLAS
project), where the project could in of itself possibly be a series of
independent applications, each with application in turn a reference to a defined
science objective achieved using a set of executables.

@@

Against this backdrop, we eschew a formal classification of application types,
however, to facilitate understanding we categorise applications into two
categories: the first is that of {\em traditional applications}, where
% traditional is  applications are {\em traditional applications}, in that 
a program (or set of programs) is developed by a user or project to try to
answer a science question (or set of related questions). Traditional application
might A single standalone computer program, a distributed application or a
complex workflow could fall under the category of a {\em traditional
  application}.  Applications in this category are: NGS (next generation
sequencing) Analytics (\S\ref{bioSilvia}), CMB (cosmic microwave background)
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial Incident
Notification and Response (\S\ref{Kees}), MODIS (moderate resolution imaging
spectroradiometer) Data Processing (\S\ref{modisKeith}), and Distributed Network
Intrusion Detection (\S\ref{detectionJon}).

Another group of applications are {\em infrastructural applications}.
Applications of this type are collaborative, in that they depend on a context
that is often agreed upon by a community, most often in terms of how data is
stored by the community and how it is accessed, as well as, more often than not,
a weak consensus on the infrastructure to be used.  This allows different
sub-applications (that we again refer to as `applications') to focus on
different stages, such as generating data, or processing data.  Answering a
science question now may involve a set of applications that need to be run in
series, perhaps in different stages, that may be run by different groups that do
not frequently interact.  This type of application includes: ATLAS/WLCG (a
toroidal large hadron collider apparatus/worldwide large hadron collider
computing grid, \S\ref{WLCGSteve}), LSST (large synoptic survey telescope,
\S\ref{astroAdam}), SOA (service-oriented architecture) Astronomy
(\S\ref{astroSOAAdam}), Sensor Network Application (\S\ref{sensorSimon}),
Climate (\S\ref{climateDan}), Interactive Exploration of Environmental Data
(\S\ref{envJon}), and Power Grids (\S\ref{power}).

In other words, multiple infrastructural applications are used together in
particular science domains.  Furthermore, infrastructural applications involve
more than just the end user (scientist), and typically require exerting control
at layers below, i.e., ``programming the system (infrastructure)'', as opposed
to a traditional applications, where end user effort and control is often
confined to a well-defined application kernel and their communication and
coordination.  Also, for some {\em traditional applications} (e.g., NGS) the
number of users of an instance of the application is typically small, i.e., the
predominant use is by the long-tail of science. In contrast, some
infrastructural applications are correlated with ``big science'' projects.

>>>>

The term ``application'' is often overloaded in literature and the computational
science discourse.  Sometimes the term is a reference to a standalone,
independent executable (e.g., a Molecular Dynamics simulation kernel such as
AMBER or Gromacs). Sometimes it refers to science problem to which
the executable is put to use (e.g., computing the free-energy of binding energy of
a drug candidate), but sometimes it is also used to reference the end-to-end
workflow (i.e., a multi-stage computational process requiring a sequence of
possibly distinct simulations and executables).  To add chaos to the confusion,
applications are sometimes also as a proxy for projects (e.g., the ATLAS
project), where the project could itself possibly be a series of
independent applications, with each application in turn a reference to a defined
science objective achieved using a set of executables.

@@

Against this backdrop, we eschew a formal classification of application types,
however, to facilitate understanding we categorize applications into two
categories: the first is that of {\em traditional applications}, where
% traditional is  applications are {\em traditional applications}, in that 
a program (or set of programs) is developed by a user or project to try to
answer a science question (or set of related questions). 
A single standalone computer program, a distributed application, or a
complex workflow could fall under the category of a {\em traditional
application}.  Applications in this paper that are in this category are: NGS (next generation
sequencing) Analytics (\S\ref{bioSilvia}), CMB (cosmic microwave background)
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial Incident
Notification and Response (\S\ref{Kees}), MODIS (moderate resolution imaging
spectroradiometer) Data Processing (\S\ref{modisKeith}), and Distributed Network
Intrusion Detection (\S\ref{detectionJon}).

Another group of applications are {\em infrastructural applications}.
Applications of this type are collaborative, in that they depend on a context
that is often agreed upon by a community, most often in terms of how data is
stored by the community and how it is accessed, as well as, more often than not,
a weak consensus on the infrastructure to be used.  This allows different
sub-applications (that we again refer to as `applications') to focus on
different stages, such as generating data, or processing data.  Answering a
science question now may involve a set of applications that need to be run in
series, perhaps in different stages, that may be run by different groups that do
not frequently interact.  In this paper, examples of this type of application includes: ATLAS/WLCG (a
toroidal large hadron collider apparatus/worldwide large hadron collider
computing grid, \S\ref{WLCGSteve}), LSST (large synoptic survey telescope,
\S\ref{astroAdam}), SOA (service-oriented architecture) Astronomy
(\S\ref{astroSOAAdam}), Sensor Network Application (\S\ref{sensorSimon}),
Climate (\S\ref{climateDan}), Interactive Exploration of Environmental Data
(\S\ref{envJon}), and Power Grids (\S\ref{power}).

In other words, multiple infrastructural applications are used together in
particular science domains.  Furthermore, infrastructural applications involve
more than just the end user (scientist), and typically require exerting control
at lower layers, i.e., ``programming the system (infrastructure)'', as opposed
to a traditional applications where end user effort and control is often
confined to a well-defined application kernel and their communication and
coordination.  Also, for some {\em traditional applications} (e.g., NGS) the
number of users of an instance of the application is typically small, i.e., the
predominant use is by the long-tail of science. In contrast, some
infrastructural applications are correlated with ``big science'' projects.
